---
title       : "Predicting child's height from parents' heights"
subtitle    : Quick and dirty Shiny app for Coursera DDP course
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]     # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## What does it do?

1. User provides his/her, as well as partner's or hypothetical partner's height and sex
2. The app uses *father.son* dataset from **UsingR** package to build a simple linear regression model
3. Child's height prediction is obtained from parents' adjusted average height and is output on the screen together with the prediction error and the accompanying plot

--- .class #id 

## Inner workings of the app

```{r, echo = FALSE, results='hide'}
library(UsingR)
library(ggplot2)
data(father.son)
```
- The linear prediction model is built using the following command:

```{r, results = 'hide'}
fit <- lm(sheight~fheight,data=father.son)
```

- This results in a model with the following coefficients:

```{r, echo = FALSE}
summary(fit)$coef
```

--- .class #id

## Inner workings of the app (contd.)
- User inputs their sex and height as well as their partner's sex and height
- Heights are input in centimeters and the prediction result is in centimeters too, while the dataset is in inches
- To provide conversion between units, two functions are implemented:

```{r, results = 'hide'}
in2cm <- function(inches) {inches/0.39370}
cm2in <- function(cm) {cm * 0.39370}
```
- Female parent's height is adjusted with multiplication by 1.08 and midheight of the two parents is calculated inside a function and then subsequently used to predict child's height with:
 
```{r, results = 'hide'}
getChildHeight <- function(lmfit, newData) {
    predict(lmfit, data.frame(fheight = c(newData)), interval = "prediction")
}
```

--- .class #id

## Example
- Suppose user is female and is 160 cm high, while their partner is male and is 180 cm high
```{r, results = 'hide', echo = FALSE}
fheight <- 160 * 1.08;
mheight <- 180;
midHeight <- mean(c(mheight,fheight))
```

- Parents' adjusted average height is then `r midHeight` cm.

```{r, results = 'hide', echo = FALSE}
childHeightPrediction <- getChildHeight(fit, cm2in(midHeight))
childHeight <- round(in2cm(childHeightPrediction[1]), 1)
stdError <- round(in2cm(childHeightPrediction[1] - childHeightPrediction[2]),1)
```

- Child's predicted height is then `r childHeight` +/- `r stdError` cm.

- If the user inputs heights such that the adjusted average height is outside of the data cloud, then the script outputs the following message:
    
    "PREDICTION UNRELIABLE, IT IS OUTSIDE OF THE DATA CLOUD!"
    
- Similarly is the user specifies same sex couple, then the script states:
    
    "Unfortunately script doesn't work for same sex couples."

---



